\Large \bf Uncovering Bugs in Distributed Storage Systems\\during Testing (not in Production!)

### 2016-01-22 evaluation.tex
- Act on Daniel Jackson's feedback on my sections. - Some other edits to evaluation.   - Move "bug found" column before "time" and "NDC" because it     affects their interpretation, i.e., whether they reflect the     standard input generator or a custom test case.   - Now that we have stars in the "bug found" column (great!), the     diamonds on the bugs are redundant, so remove them.   - Replace star with diamond because diamond is easier to     distinguish from check and X.   - Reword caption (see included comment).

\setlength{\tabcolsep}{.68em}
\begin{table*}[t]
\small
\centering
\begin{tabular}{rl rrr rrr}
\centering
\input{experiments/tables/results.tex}
\end{tabular}
\caption{Results from running the \psharp random and priority-based systematic testing schedulers for 100,000 executions. We report: time in seconds to find a bug (Time to Bug); number of nondeterministic choices made in the first execution that a bug was found (\#NDC); and if a bug was found (BF?) or not (\cmark\xspace denotes that a bug was found, \smark\xspace denotes that a bug was found only using a custom test harness, and \xmark\xspace denotes that a bug could not be found).}
\label{tab:testing}
%\vspace{-2mm}
\end{table*}

@@

For the MigratingTable case study, we evaluated the \psharp test harness of \S\ref{sec:migrating} on eleven bugs, including eight \emph{organic} bugs that actually occurred in development and three \emph{notional} bugs (denoted by $*$), which are other code changes that we deemed interesting ways of making the system incorrect. The harness found seven of the organic bugs, which are listed first in Table~\ref{tab:testing}. The remaining four bugs (denoted by $\diamond$) were not caught with our default test harness in the 100,000 executions (a process that required less than 30 minutes). We believe this is because the inputs and schedules that trigger them are too rare in the used distribution. To confirm this, we wrote a custom test case for each bug with a specific input that triggers it and were able to quickly reproduce the bugs;
% Be a little bit more transparent about the fact that the $\diamond$ rows do not
% reflect results achievable via our main testing approach. ~ Matt 2015-12-31
the table shows the results of these runs.
Note that the random scheduler managed to only find seven of the MigratingTable bugs; we had to use the priority-based scheduler to discover the remaining bugs.

The QueryStreamedBackUpNewStream bug in MigratingTable, that was found using \psharp, stands out because it reflects a type of oversight that can easily occur as systems evolve. This bug is in the implementation of a \emph{streaming read} from the virtual table, which should return a stream of all rows in the table sorted by key. The essential implementation idea is to perform streaming reads from both backend tables and merge the results. According to the \texttt{IChainTable} specification, each row read from a stream may reflect the state of the table at any time between when the stream was started and the row was read. The developers sketched a proof that the merging process would preserve this property as long as the migrator was only copying rows from the old table to the new table. But when support was added for the migrator to delete the rows from the old table after copying, it became possible for the backend streams to see the deletion of a row from the old table but not its insertion into the new table, even though the insertion happens first, and the row would be missed. \psharp managed to discover this bug in a matter of seconds. The MigratingTable developers spent just 10 minutes analyzing the trace to diagnose what was happening; although, this was after days of experience analyzing traces and after they had extended the logging of \psharp with additional trace information specific to MigratingTable. \psharp only outputs trace information related to its communicating state machines, but the trace information is easy to extend, and we extended trace information in a custom fashion for each of our case studies.

>>>>

\setlength{\tabcolsep}{.68em}
\begin{table*}[t]
\small
\centering
\begin{tabular}{rl rrr rrr}
\centering
\input{experiments/tables/results.tex}
\end{tabular}
\caption{Results from running the \psharp random and priority-based systematic testing schedulers for 100,000 executions. We report: whether the bug was found (BF?) (\cmark\xspace means it was found, $\diamond$\xspace means it was found only using a custom test case, and \xmark\xspace means it was not found); time in seconds to find the bug (Time to Bug); and number of nondeterministic choices made in the first execution that found the bug (\#NDC).}
% Changing ``a bug'' to ``the bug'' makes this a lot easier to read.  The risk
% of this change is that we're claiming that all failures of the P# test in each
% row represent the listed bug, rather than just ``a bug''.  I went to enough
% trouble to verify this for MigratingTable that I'm comfortable making this
% claim, although I don't have complete proof.  In any case, I don't think this
% is a distinction we can reasonably hope to get across to the reader under our
% space constraint. ~ Matt 2016-01-22
\label{tab:testing}
%\vspace{-2mm}
\end{table*}

@@

For the MigratingTable case study, we evaluated the \psharp test harness of \S\ref{sec:migrating} on eleven bugs, including eight \emph{organic} bugs that actually occurred in development and three \emph{notional} bugs (denoted by $*$), which are other code changes that we deemed interesting ways of making the system incorrect. The harness found seven of the organic bugs, which are listed first in Table~\ref{tab:testing}. The remaining four bugs (marked $\diamond$) were not caught with our default test harness in the 100,000 executions (a process that required less than 30 minutes). We believe this is because the inputs and schedules that trigger them are too rare in the used distribution. To confirm this, we wrote a custom test case for each bug with a specific input that triggers it and were able to quickly reproduce the bugs;
% Be a little bit more transparent about the fact that the $\diamond$ rows do not
% reflect results achievable via our main testing approach. ~ Matt 2015-12-31
the table shows the results of these runs.
Note that the random scheduler only managed to trigger seven of the MigratingTable bugs; we had to use the priority-based scheduler to trigger the remaining bugs.
% The words ``find'' and ``discover'' are misleading for the diamond bugs.
% ~ Matt 2016-01-21

The QueryStreamedBackUpNewStream bug in MigratingTable, which was found using \psharp, stands out because it reflects a type of oversight that can easily occur as systems evolve. This bug is in the implementation of a \emph{streaming read} from the virtual table, which should return a stream of all rows in the table sorted by key. The essential implementation idea is to perform streaming reads from both backend tables and merge the results. According to the \texttt{IChainTable} specification, each row read from a stream may reflect the state of the table at any time between when the stream was started and the row was read. The developers sketched a proof that the merging process would preserve this property as long as the migrator was only copying rows from the old table to the new table. But when support was added for the migrator to delete the rows from the old table after copying, it became possible for the backend streams to see the deletion of a row from the old table but not its insertion into the new table, even though the insertion happens first, and the row would be missed. The \psharp test discovered this bug in a matter of seconds. The MigratingTable developers spent just 10 minutes analyzing the trace to diagnose what was happening, although admittedly, this was after they added MigratingTable-specific trace information and had several days of experience analyzing traces. (Out of the box, \psharp traces include only machine- and event-level information, but it is easy to add application-specific information, and we did so in all of our case studies.)

### 2016-01-22 migrating.tex
updates towards camera ready
- Act on Daniel Jackson's feedback on my sections. - Some other edits to evaluation.   - Move "bug found" column before "time" and "NDC" because it     affects their interpretation, i.e., whether they reflect the     standard input generator or a custom test case.   - Now that we have stars in the "bug found" column (great!), the     diamonds on the bugs are redundant, so remove them.   - Replace star with diamond because diamond is easier to     distinguish from check and X.   - Reword caption (see included comment).

% Remove the \emph on ``transparently'' and ``while'' again.  Pantazis, I'm
% imagining you were trying to emphasize what's interesting about the problem,
% but I think suggesting that this idea is new to the reader is insulting.
% ~ Matt 2015-12-31
Live Table Migration (MigratingTable) is a library designed to transparently migrate a data set between tables in the Azure storage service while an application is accessing this data set. This case study is particularly interesting because MigratingTable was \emph{co-developed} with its \psharp test harness. This co-development process resulted in the discovery of several tricky bugs before the system went into production (see \S\ref{sec:eval}). The MigratingTable testing effort differs from the vNext case study: the vNext developers focused on a single liveness property, while the MigratingTable test harness checks complete compliance with an interface (safety) specification.

MigratingTable provides a \emph{virtual table} (VT) that has a similar interface
to an ordinary Azure table. This interface is
named \texttt{IChainTable}. Each VT is backed by a pair
of \emph{old} and \emph{new} \emph{backend} tables (BTs), both of which
implement \texttt{IChainTable}. A \emph{migrator} job is responsible
for moving all data from the old table to the new table in the
background. Meanwhile, each \emph{input} read and write operation
issued to the VT is implemented via a sequence
of reads and writes on the BTs according to a protocol,
which intends to guarantee that when multiple applications issue input
operations to different MigratingTable instances with the same BTs, the output of these operations complies with the specification
of \texttt{IChainTable} for the combined \emph{input history}, as if
all the operations were performed on a single table. The goal of
using \psharp was to systematically test this property.

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{img/modeled_migration}
\caption{The \psharp test harness of MigratingTable (each box with a dotted line represents one \psharp machine).}
\label{fig:mockedmigration}
\end{figure}

There are two main challenges behind testing MigratingTable: (i) the system is highly concurrent; and (ii) the input operations accept many parameters that affect the behavior in different ways. The developers could have chosen specific input histories to test, but they were not confident that they would cover the combinations of parameters that might trigger bugs. Instead, the developers used the \psharp nondeterministic choice generation APIs to sample from a distribution over input histories that they defined, which allowed all of the relevant parameters to vary within certain limits. They created a reference implementation of the \texttt{IChainTable} specification, called \texttt{SpecTable}, and compared the output of MigratingTable on the chosen input history to that of a \texttt{SpecTable} instance called the \emph{reference table}. The \texttt{SpecTable} implementation was reused for the BTs of MigratingTable.
% I'm removing the claim that each bug is caught with positive probability per
% run because it's too awkward to explain here and it's nothing specific to
% MigratingTable.  Also, we can't quite appeal to the small scope hypothesis in
% its original form because I hand-picked which parameters to vary and what
% limits to use.  The value of the input generator should be clear enough from
% the above. ~ Matt 2015-12-31

Figure~\ref{fig:mockedmigration} shows the \psharp test harness of MigratingTable. The test harness consists of a \texttt{Tables} machine, which contains the BTs and the reference table, and serializes all operations on these tables; a set of \texttt{Service} machines that contain identically configured instances of MigratingTable; and a \texttt{Migrator} machine that performs the background migration. Each \texttt{Service} machine issues a random sequence of input operations to its MigratingTable, which sends the necessary operations to the BTs via \psharp events. The developers instrumented MigratingTable to report the \emph{linearization point} of each input operation, i.e., the time at which it takes effect on the single VT, so the test harness could perform the operation on the reference table at the same time. More precisely, after processing each backend operation, the \texttt{Tables} machine enters a \psharp state that blocks all further work until MigratingTable reports whether the backend operation was the linearization point and, if so, the input operation has been completed on the reference table. This way, the rest of the system never observes the reference table to be out of sync with the VT.
% Note: If the tables were observed to be out of sync, it could cause both false
% positives and false negatives. ~ Matt 2015-12-31

For further information about this case study, we refer the reader to the source code repository~\cite{migratingtable-src}.

>>>>

% Remove the \emph on ``transparently'' and ``while'' again.  Pantazis, I'm
% imagining you were trying to emphasize what's interesting about the problem,
% but I think suggesting that this idea is new to the reader is insulting.
% ~ Matt 2015-12-31
%
% Delete ``safety'': the MigratingTable harness does have a crude liveness
% check. ~ Matt 2016-01-20
Live Table Migration (MigratingTable) is a library designed to transparently migrate a key-value data set between two Microsoft Azure tables~\cite{azure-table} (called the \emph{old} table and the \emph{new} table, or collectively the \emph{backend} tables) while an application is accessing this data set. The MigratingTable testing effort differs from the vNext effort in two significant ways: the \psharp test harness was developed along with the system rather than later, and it checks complete compliance with an interface specification rather than just a single liveness property. Indeed, the \psharp test caught bugs throughout the development process (see \S\ref{sec:eval}), and the developers did not bother with any other correctness tests.

During a migration, each application process creates its own MigratingTable instance (MT) referring to the same backend tables (BTs) and performs all data access via the MT. (The MT provides an interface named \texttt{IChainTable} similar to that of an Azure table, and it assumes the BTs provide the same interface via an adapter.) A \emph{migrator} job moves the data in the background. In the meantime, each \emph{logical} read or write operation issued to an MT is implemented via a sequence of \emph{backend} operations on the BTs according to a custom protocol. The protocol is designed to guarantee that the output of the logical operations complies with the \texttt{IChainTable} specification as if all the operations were performed on a single \emph{virtual table} (VT). The goal of using \psharp was to systematically test this property.

There are two main challenges behind testing MigratingTable: (i) the system is highly concurrent; and (ii) the logical operations accept many parameters that affect the behavior in different ways. The developers could have chosen specific input sequences, but they were not confident that these sequences would cover the combinations of parameters that might trigger bugs. Instead, they used the \psharp nondeterministic choice API to choose all of the parameters independently within certain limits. They issued the same operations to the MTs and to a \emph{reference table} (RT) running a reference implementation of the \texttt{IChainTable} specification, and compared the output. This reference implementation was reused for the BTs, since the goal was not to test the real Azure tables.
% I'm removing the claim that each bug is caught with positive probability per
% run because it's too awkward to explain here and it's nothing specific to
% MigratingTable.  Also, we can't quite appeal to the small scope hypothesis in
% its original form because I hand-picked which parameters to vary and what
% limits to use.  The value of the input generator should be clear enough from
% the above. ~ Matt 2015-12-31

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{img/modeled_migration}
\vspace{-6mm}
\caption{The \psharp test environment of MigratingTable (each box with a dotted line represents one \psharp machine).}
\label{fig:mockedmigration}
\vspace{-2mm}
\end{figure}

The complete test environment is shown in Figure~\ref{fig:mockedmigration}. It consists of a \texttt{Tables} machine, which contains the BTs and RT, and serializes all operations on these tables; a set of \texttt{Service} machines that contain identically configured MTs; and a \texttt{Migrator} machine that performs the background migration. Each \texttt{Service} machine issues a random sequence of logical operations to its MT, which performs the backend operations on the BTs via \psharp events. The developers instrumented the MTs to report the \emph{linearization point} of each logical operation, i.e., the time at which it takes effect on the VT, so the test harness could perform the operation on the RT at the same time. (More precisely, after processing each backend operation, the \texttt{Tables} machine enters a \psharp state that blocks all further work until the MT reports whether the backend operation was the linearization point and, if so, the logical operation has been completed on the RT. This way, the rest of the system never observes the RT to be out of sync with the VT.)
% In my reading, the parentheses make it clear that we're amending the previous
% sentence; without them, it looks more like a contradiction. ~ Matt 2016-01-21
% Note: If the tables were observed to be out of sync, it could cause both false
% positives and false negatives. ~ Matt 2015-12-31
% Join paragraphs to save space
For further implementation details of MigratingTable and its test harness, we refer the reader to the source code repository~\cite{migratingtable-src}.
